\documentclass[../main.tex]{subfiles}

\begin{document}
	
\section{IoT Software Update Manager Prototype}

We provide an open-source implementation of the SUIT workflow which we integrated in the main branch o RIOT~\cite{RIOT-suit}, a common operating system for low-power IoT devices~\cite{baccelli2018riot} which we used as base for our  microcontainer prototypes.

Our update manager prototype is generic in the sense that it enables both to manage the secure update of:
\begin{itemize}
    \item Software modules (i.e. OTA microcontainer update);
    \item Firmware (FOTA updates)
\end{itemize}

The microcontainer update demo shows the integration between the secure update
manager and the microcontainer runtime environment.
The secure update manager leverages SUIT as described above to deliver the
application payload.
The microcontainer runtime uses Femto-Containers to run the applications.

The demonstrator uses two in-memory slots that can be updated with new payloads.
Both of these in-memory slots are hooked up to the Femto-Container CoAP
endpoints. These act as a demonstrator for the SUIT capabilities together with
Femto-Containers. After an application is loaded in either of the available
slots, the CoAP endpoint can be used to trigger the Femto-Container application
and the CoAP response will contain the resulting return value from the
Femto-Container execution.

The in-memory slots are updated using the SUIT workflow in RIOT, as described in
the \texttt{README.md} file present with the deliverable.

\end{document}
